Deep Learning for 3D Brain Activity Scans

### 2019-11-11 paper/submission.tex
Updating paper, and some misc code edits I forget about

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}
\begin{document}

%%%%%%%%% TITLE
\title{\LaTeX\ Author Guidelines for CVPR Proceedings}

\author{First Author\\
Institution1\\
Institution1 address\\
{\tt\small firstauthor@i1.org}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Second Author\\
Institution2\\
First line of institution2 address\\
{\tt\small secondauthor@i2.org}
}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
   The ABSTRACT is to be in fully-justified italicized text, at the top
   of the left-hand column, below the author and affiliation
   information. Use the word ``Abstract'' as the title, in 12-point
   Times, boldface type, centered relative to the column, initially
   capitalized. The abstract is to be in 10-point, single-spaced type.
   Leave two blank lines after the Abstract, then begin the main text.
   Look at previous CVPR abstracts to get a feel for style and length.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

Please follow the steps outlined below when submitting your manuscript to
the IEEE Computer Society Press.  This style guide now has several
important modifications (for example, you are no longer warned against the
use of sticky tape to attach your artwork to the paper), so all authors
should read this new version.

%-------------------------------------------------------------------------
\subsection{Language}

All manuscripts must be in English.

\subsection{Dual submission}

Please refer to the author guidelines on the CVPR 2018 web page for a
discussion of the policy on dual submissions.

\subsection{Paper length}
Papers, excluding the references section,
must be no longer than eight pages in length. The references section
will not be included in the page count, and there is no limit on the
length of the references section. For example, a paper of eight pages
with two pages of references would have a total length of 10 pages.
{\bf There will be no extra page charges for CVPR 2018.}

Overlength papers will simply not be reviewed.  This includes papers
where the margins and formatting are deemed to have been significantly
altered from those laid down by this style guide.  Note that this
\LaTeX\ guide already sets figure captions and references in a smaller font.
The reason such papers will not be reviewed is that there is no provision for
supervised revisions of manuscripts.  The reviewing process cannot determine
the suitability of the paper for presentation in eight pages if it is
reviewed in eleven.

%-------------------------------------------------------------------------
\subsection{The ruler}
The \LaTeX\ style defines a printed ruler which should be present in the
version submitted for review.  The ruler is provided in order that
reviewers may comment on particular lines in the paper without
circumlocution.  If you are preparing a document using a non-\LaTeX\
document preparation system, please arrange for an equivalent ruler to
appear on the final output pages.  The presence or absence of the ruler
should not change the appearance of any other content on the page.  The
camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
the \verb'\cvprfinalcopy' command in the document preamble.)  Reviewers:
note that the ruler measurements do not align well with lines in the paper
--- this turns out to be very difficult to do well when the paper contains
many figures and equations, and, when done, looks ugly.  Just use fractional
references (e.g.\ this line is $095.5$), although in most cases one would
expect that the approximate location will be adequate.

\subsection{Mathematics}

Please number all of your sections and displayed equations.  It is
important for readers to be able to refer to any particular equation.  Just
because you didn't refer to it in the text doesn't mean some future reader
might not need to refer to it.  It is cumbersome to have to use
circumlocutions like ``the equation second from the top of page 3 column
1''.  (Note that the ruler will not be present in the final copy, so is not
an alternative to equation numbers).  All authors will benefit from reading
Mermin's description of how to write mathematics:
\url{http://www.pamitc.org/documents/mermin.pdf}.

\subsection{Blind review}

Many authors misunderstand the concept of anonymizing for blind
review.  Blind review does not mean that one must remove
citations to one's own work---in fact it is often impossible to
review a paper unless the previous citations are known and
available.

Blind review means that you do not use the words ``my'' or ``our''
when citing previous work.  That is all.  (But see below for
techreports.)

Saying ``this builds on the work of Lucy Smith [1]'' does not say
that you are Lucy Smith; it says that you are building on her
work.  If you are Smith and Jones, do not say ``as we show in
[7]'', say ``as Smith and Jones show in [7]'' and at the end of the
paper, include reference 7 as you would any other cited work.

An example of a bad paper just asking to be rejected:
\begin{quote}
\begin{center}
    An analysis of the frobnicatable foo filter.
\end{center}

   In this paper we present a performance analysis of our
   previous paper [1], and show it to be inferior to all
   previously known methods.  Why the previous paper was
   accepted without this analysis is beyond me.

   [1] Removed for blind review
\end{quote}

An example of an acceptable paper:

\begin{quote}
\begin{center}
     An analysis of the frobnicatable foo filter.
\end{center}

   In this paper we present a performance analysis of the
   paper of Smith \etal [1], and show it to be inferior to
   all previously known methods.  Why the previous paper
   was accepted without this analysis is beyond me.

   [1] Smith, L and Jones, C. ``The frobnicatable foo
   filter, a fundamental contribution to human knowledge''.
   Nature 381(12), 1-213.
\end{quote}

If you are making a submission to another conference at the same time,
which covers similar or overlapping material, you may need to refer to that
submission in order to explain the differences, just as you would if you
had previously published related work.  In such cases, include the
anonymized parallel submission~\cite{Authors14} as additional material and
cite it as
\begin{quote}
[1] Authors. ``The frobnicatable foo filter'', F\&G 2014 Submission ID 324,
Supplied as additional material {\tt fg324.pdf}.
\end{quote}

Finally, you may feel you need to tell the reader that more details can be
found elsewhere, and refer them to a technical report.  For conference
submissions, the paper must stand on its own, and not {\em require} the
reviewer to go to a techreport for further details.  Thus, you may say in
the body of the paper ``further details may be found
in~\cite{Authors14b}''.  Then submit the techreport as additional material.
Again, you may not assume the reviewers will read this material.

Sometimes your paper is about a problem which you tested using a tool which
is widely known to be restricted to a single institution.  For example,
let's say it's 1969, you have solved a key problem on the Apollo lander,
and you believe that the CVPR70 audience would like to hear about your
solution.  The work is a development of your celebrated 1968 paper entitled
``Zero-g frobnication: How being the only people in the world with access to
the Apollo lander source code makes us a wow at parties'', by Zeus \etal.

You can handle this paper like any other.  Don't write ``We show how to
improve our previous work [Anonymous, 1968].  This time we tested the
algorithm on a lunar lander [name of lander removed for blind review]''.
That would be silly, and would immediately identify the authors. Instead
write the following:
\begin{quotation}
\noindent
   We describe a system for zero-g frobnication.  This
   system is new because it handles the following cases:
   A, B.  Previous systems [Zeus et al. 1968] didn't
   handle case B properly.  Ours handles it by including
   a foo term in the bar integral.

   ...

   The proposed system was integrated with the Apollo
   lunar lander, and went all the way to the moon, don't
   you know.  It displayed the following behaviours
   which show how well we solved cases A and B: ...
\end{quotation}
As you can see, the above text follows standard scientific convention,
reads better than the first version, and does not explicitly name you as
the authors.  A reviewer might think it likely that the new paper was
written by Zeus \etal, but cannot make any decision based on that guess.
He or she would have to be sure that no other authors could have been
contracted to solve problem B.
\medskip

\noindent
FAQ\medskip\\
{\bf Q:} Are acknowledgements OK?\\
{\bf A:} No.  Leave them for the final copy.\medskip\\
{\bf Q:} How do I cite my results reported in open challenges?
{\bf A:} To conform with the double blind review policy, you can report results of other challenge participants together with your results in your paper. For your results, however, you should not identify yourself and should not mention your participation in the challenge. Instead present your results referring to the method proposed in your paper and draw conclusions based on the experimental comparison to other results.\medskip\\

\begin{figure}[t]
\begin{center}
\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   %\includegraphics[width=0.8\linewidth]{egfigure.eps}
\end{center}
   \caption{Example of caption.  It is set in Roman so that mathematics
   (always set in Roman: $B \sin A = A \sin B$) may be included without an
   ugly clash.}
\label{fig:long}
\label{fig:onecol}
\end{figure}

\subsection{Miscellaneous}

\noindent
Compare the following:\\
\begin{tabular}{ll}
 \verb'$conf_a$' &  $conf_a$ \\
 \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
\end{tabular}\\
See The \TeX book, p165.

The space after \eg, meaning ``for example'', should not be a
sentence-ending space. So \eg is correct, {\em e.g.} is not.  The provided
\verb'\eg' macro takes care of this.

When citing a multi-author paper, you may save space by using ``et alia'',
shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word.)
However, use it only when there are three or more authors.  Thus, the
following is correct: ``
   Frobnication has been trendy lately.
   It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
   Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''

This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...''
because reference~\cite{Alpher03} has just two authors.  If you use the
\verb'\etal' macro provided, then you need not worry about double periods
when used at the end of a sentence as in Alpher \etal.

For this citation style, keep multiple citations in numerical (not
chronological) order, so prefer \cite{Alpher03,Alpher02,Authors14} to
\cite{Alpher02,Alpher03,Authors14}.

\begin{figure*}
\begin{center}
\fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
\end{center}
   \caption{Example of a short caption, which should be centered.}
\label{fig:short}
\end{figure*}

%------------------------------------------------------------------------
\section{Formatting your paper}

All text must be in a two-column format. The total allowable width of the
text area is $6\frac78$ inches (17.5 cm) wide by $8\frac78$ inches (22.54
cm) high. Columns are to be $3\frac14$ inches (8.25 cm) wide, with a
$\frac{5}{16}$ inch (0.8 cm) space between them. The main title (on the
first page) should begin 1.0 inch (2.54 cm) from the top edge of the
page. The second and following pages should begin 1.0 inch (2.54 cm) from
the top edge. On all pages, the bottom margin should be 1-1/8 inches (2.86
cm) from the bottom edge of the page for $8.5 \times 11$-inch paper; for A4
paper, approximately 1-5/8 inches (4.13 cm) from the bottom edge of the
page.

%-------------------------------------------------------------------------
\subsection{Margins and page numbering}

All printed material, including text, illustrations, and charts, must be kept
within a print area 6-7/8 inches (17.5 cm) wide by 8-7/8 inches (22.54 cm)
high.
Page numbers should be in footer with page numbers, centered and .75
inches from the bottom of the page and make it start at the correct page
number rather than the 4321 in the example.  To do this fine the line (around
line 23)
\begin{verbatim}
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{4321}
\end{verbatim}
where the number 4321 is your assigned starting page.

Make sure the first page is numbered by commenting out the first page being
empty on line 46
\begin{verbatim}
%\thispagestyle{empty}
\end{verbatim}

%-------------------------------------------------------------------------
\subsection{Type-style and fonts}

Wherever Times is specified, Times Roman may also be used. If neither is
available on your word processor, please use the font closest in
appearance to Times to which you have access.

MAIN TITLE. Center the title 1-3/8 inches (3.49 cm) from the top edge of
the first page. The title should be in Times 14-point, boldface type.
Capitalize the first letter of nouns, pronouns, verbs, adjectives, and
adverbs; do not capitalize articles, coordinate conjunctions, or
prepositions (unless the title begins with such a word). Leave two blank
lines after the title.

AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
and printed in Times 12-point, non-boldface type. This information is to
be followed by two blank lines.

The ABSTRACT and MAIN TEXT are to be in a two-column format.

MAIN TEXT. Type main text in 10-point Times, single-spaced. Do NOT use
double-spacing. All paragraphs should be indented 1 pica (approx. 1/6
inch or 0.422 cm). Make sure your text is fully justified---that is,
flush left and flush right. Please do not place any additional blank
lines between paragraphs.

Figure and table captions should be 9-point Roman type as in
Figures~\ref{fig:onecol} and~\ref{fig:short}.  Short captions should be centred.

\noindent Callouts should be 9-point Helvetica, non-boldface type.
Initially capitalize only the first word of section titles and first-,
second-, and third-order headings.

FIRST-ORDER HEADINGS. (For example, {\large \bf 1. Introduction})
should be Times 12-point boldface, initially capitalized, flush left,
with one blank line before, and one blank line after.

SECOND-ORDER HEADINGS. (For example, { \bf 1.1. Database elements})
should be Times 11-point boldface, initially capitalized, flush left,
with one blank line before, and one after. If you require a third-order
heading (we discourage it), use 10-point Times, boldface, initially
capitalized, flush left, preceded by one blank line, followed by a period
and your text on the same line.

%-------------------------------------------------------------------------
\subsection{Footnotes}

Please use footnotes\footnote {This is what a footnote looks like.  It
often distracts the reader from the main flow of the argument.} sparingly.
Indeed, try to avoid footnotes altogether and include necessary peripheral
observations in
the text (within parentheses, if you prefer, as in this sentence).  If you
wish to use a footnote, place it at the bottom of the column on the page on
which it is referenced. Use Times 8-point type, single-spaced.

%-------------------------------------------------------------------------
\subsection{References}

List and number all bibliographical references in 9-point Times,
single-spaced, at the end of your paper. When referenced in the text,
enclose the citation number in square brackets, for
example~\cite{Authors14}.  Where appropriate, include the name(s) of
editors of referenced books.

\begin{table}
\begin{center}
\begin{tabular}{|l|c|}
\hline
Method & Frobnability \\
\hline\hline
Theirs & Frumpy \\
Yours & Frobbly \\
Ours & Makes one's heart Frob\\
\hline
\end{tabular}
\end{center}
\caption{Results.   Ours is better.}
\end{table}

%-------------------------------------------------------------------------
\subsection{Illustrations, graphs, and photographs}

All graphics should be centered.  Please ensure that any point you wish to
make is resolvable in a printed copy of the paper.  Resize fonts in figures
to match the font in the body text, and choose line widths which render
effectively in print.  Many readers (and reviewers), even of an electronic
copy, will choose to print your paper in order to read it.  You cannot
insist that they do otherwise, and therefore must not assume that they can
zoom in to see tiny details on a graphic.

When placing figures in \LaTeX, it's almost always best to use
\verb+\includegraphics+, and to specify the  figure width as a multiple of
the line width as in the example below
{\small\begin{verbatim}
   \usepackage[dvips]{graphicx} ...
   \includegraphics[width=0.8\linewidth]
                   {myfile.eps}
\end{verbatim}
}

%-------------------------------------------------------------------------
\subsection{Color}

Please refer to the author guidelines on the CVPR 2018 web page for a discussion
of the use of color in your document.

%------------------------------------------------------------------------
\section{Final copy}

You must include your signed IEEE copyright release form when you submit
your finished paper. We MUST have this form before your paper can be
published in the proceedings.

Please direct any questions to the production editor in charge of these
proceedings at the IEEE Computer Society Press: Phone (714) 821-8380, or
Fax (714) 761-1784.

{\small
\bibliographystyle{ieee}
\bibliography{egbib}
}

>>>>

% Pages are numbered in submission mode, and unnumbered in camera-ready
%\ifcvprfinal\pagestyle{empty}\fi
\setcounter{page}{1}
\begin{document}

%-------------------------------------------------------------------------
\title{Deep Learning for 3D Brain Activity Scans}

\author{Evan Albright\\
University of Massachusetts Amherst\\
140 Governors Dr, Amherst, MA 01002\\
{\tt\small ealbright@umass.edu}
}

@@

%-------------------------------------------------------------------------
\begin{abstract}
   Abstract placeholder
\end{abstract}

% ----------------------------------------------------------------------
% INTRODUCTION
% ----------------------------------------------------------------------
\section{Introduction}\label{sec:introduction}
Neuroimaging presents fertile grounds for deep learning applications, with myriad interesting real world benefits
for mental health research and disease diagnosis.
High dimensional, massive datasets, with intertwining patterns of activity and structure in three or four
dimensions, are a lot to manage by hand.

Classification of activity and structures should come quite naturally with deep learning practices.
Previous works include the development of Restricted Boltzmann Machines, and Deep Belief Networks~\cite{plis2014deep},
yet often only as a proof of concept.
Others are working now to try and find universal architectures for most
neuro-imaging tasks~\cite{henschel2019fastsurfer}, though it seems this is still far from a solved
application with satisfactory performances across the board.

Of more interest to myself, deep learning networks in their proven capacity to segment images based on what entities
they identify in classification could surely be adopted to highlight what patterns of activity have informed their
own decision making and perhaps clue researchers in to overlooked details.

My ambition is to create a highly accurate 3d convolutional network to first classify, then segment brain scans to
help with my specific application of neuroimaging.

I may also attempt to explore feature analysis and reduction which has been shown to
help~\cite{shi2018feature} with deep learning tasks, if only to manage the challenging
dimensionality of their 3d inputs.

% ----------------------------------------------------------------------
% PROBLEM STATEMENT
% ----------------------------------------------------------------------
\section{Problem Statement}\label{sec:problem-statement}
The dataset to be used is a set of neuro-imaging scans from a cohort of 30 healthy, young women in the
Netherlands circa 2013 who were presented with various grids of images featuring
either neutral entities, (e.g.\ office chairs, windows, jackets) or foods~\cite{smeets2013allured}.
The foods were specifically types of calorically dense, tempting foods, which the individuals viewing would likely
have formed restraint with given their success in maintaining healthy body mass indices.
The experiment was meant to highlight brain structures roles in anticipated rewards from consuming delicious foods,
balanced against the hypothesized activation of restraints.
In neuro-imaging, this task is known as passive image viewing.

Given the quite important role of food and willpower in our lives, the activation patterns are likely to be powerful
and distinguishable from neutral objects which we haven't evolved to be nearly as concerned with.
This combination gives me high hopes of success in this specific application of learning to classify the viewed
objects through brain activations of the subjects.

I hope given the clarity in patterns in the brain scans, that I can develop a neural network to highlight what
features and locations in the brain scans have been helpful in informing the decisions of classification, in the
hope that such methods could be adapted to harder tasks in the future.

The dataset can be found from the excellent
OpenNeuro project here: \url{https://openneuro.org/datasets/ds000157/versions/00001}
It consists of a survey detailing the relationship of each subject with food and diet taken after the experimental
proceedings, and the core data of two modalities of brain imaging: a single high resolution T1w brain scan,
and time series of bold scans of about 370-375 scans, which encompasses the approximately 10 minute experiment
in its entirety.
The bold scans are quite sizeable given the number of them, however they are a relatively low resolution, consisting
of 64 sagital by 64 coronal by 30 axial scans.

You may review the full experimental paper as well here: \url{http://www.pamitc.org/documents/mermin.pdf}.

You may find this project repository here: \url{https://github.com/e-m-albright/CS682}
Reading the paper is critical to understanding the format of the dataset, however I hope I will provide enough
abstraction in the project repository for you to be freed of that challenge should you wish to have a look around.

% ----------------------------------------------------------------------
% TECHNICAL APPROACH
% ----------------------------------------------------------------------
\section{Technical Approach}\label{sec:technical-approach}
I intend to explore the efficacy of 3d convolutional networks which have been shown to have a slight performance
edge on 2d convolutions~\cite{payan2015predicting}.
Given the nature of the input images being 3d brain scans this makes intuitive sense and could allow for much more
interesting patterns of brain activity and structuring being recognized.

As for the deeper architecture, I'm hoping to imitate some of the more successful networks like VGG or ResNet,
however I wouldn't be surprised if they don't adapt well to this particular application, especially given how
interesting and important it could be to keep the input data in a three dimensional format.

It is my hope that I can also learn from other developments of object detection to see what regions in the
brain become active for a particular classification, like Faster R-CNN

\subsection{Novelty}\label{subsec:novelty}

I have not seen any application of classification or object segmentation to neuroimaging in 3d
on brain activity patterns.
I have seen some works which I hope to leverage with similar aims like that of structural identification
for tumors, in two dimensions~\cite{akkus2017deep}.
However at the very least I hope to produce a very successful network on my chosen problem first and foremost rather
than assume I am the only person to ever think of something as intuitive as bringing neural networks to bear
on neuro-imaging tasks.

I hope to find a methodology of first classifying correctly the brain scans, improve accuracy as much as possible,
and adapt the approach into tools for identifying important regions of activity in the
brain which are being utilized for classification.

I am not chiefly focused on coping with the challenges of memory but it may become too important to overlook.
Each dataset input is a $64 \times 64 \times 30$ slice of floats, for which we've got
about $375 scans \times 30 subjects$ total.
In my work on preliminary results I have already been confronted by the limitations of my current machine and have
been trying working around the issue until I am ready to attempt to resolve it.
This may be an unavoidable area to simultaneously try to innovate for the field of
neuro-imaging analysis with deep learning

% ----------------------------------------------------------------------
% INTERMEDIATE RESULTS
% ----------------------------------------------------------------------
\section{Intermediate Results}\label{sec:intermediate-results}
At the date of the milestone, I have shown that my experimental set-up is complete to begin
iterating towards the novel contributions aimed for by this project.

I am generally happy with how things are going, and feel optimistic that I will construct a successful neural
network for classification on my chosen dataset incorporating my own contributions to the process.

I hope my goal for region highlighting is also feasible in the time frame with only myself on the task, but I am
excited nevertheless to try.

\subsection{Data}\label{subsec:data}
I have become familiar with the chosen dataset, and created a machine learning friendly format
based off of the data.

I have inferred classification labels from the experimental events timings and the estimated time
of each complete scan in the dataset.

\subsection{Sanity Checks}\label{subsec:sanity-checks}
Using the data, I have implemented a very simple, yet effective 80\% accurate SVM classifier to prove the data
should be ready for use with my chosen deep learning methodologies.

I'm now confident in the classifications of each of the scans (about 375 per subject, for 30 subjects)

\subsection{Deep Learning \& Future Works}\label{subsec:deep-learning}
I have tried feeding the data into very basic neural networks with mixed success, and am in the active process of
troubleshooting.
Once I have matched performance with the SVM using an implementation of a reasonable neural network using PyTorch, I
will begin the process of looking for refinements and innovations alluded to in previous sections, as well as
transitioning the task to object detection if I can.

% ----------------------------------------------------------------------
% END
% ----------------------------------------------------------------------
{\small
\bibliographystyle{ieee}
\bibliography{project}
}
